---
title: "p8105_hw6_sl4658"
author: "Simin Ling"
date: "12/3/2020"
output: github_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(p8105.datasets)
```


## Problem 1
Create a city_state variable and a binary variable indicating whether the homicide is solved.
```{r, message=FALSE}
homicide_df = 
  read_csv("./data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    victim_race = as.factor(victim_race),
    victim_sex = as.factor(victim_sex),
    resolved = case_when(
      disposition  == "Closed without arrest" ~ 0,
      disposition  == "Open/No arrest" ~ 0,
      disposition  == "Closed by arrest" ~ 1)
  ) %>% 
  filter(
    city_state != "Tulsa, AL",
    victim_race %in% c("White", "Black")
    ) %>% 
  select(city_state, resolved, victim_age, victim_race, victim_sex)
```


Create a dataframe for Baltimore, MD
```{r}
baltimore_df = 
  homicide_df %>%
  filter(city_state == "Baltimore, MD")

head(baltimore_df)
```


Fit the logistic regression for Baltimore, MD
```{r}
glm(resolved ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>%
  broom::tidy() %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>%
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```


Run glm for each of the cities in dataset
```{r}
models_results_df = 
  homicide_df %>%
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolved ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())), 
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI"))
```


Create a plot that shows the estimated ORs and CIs for each city. 
```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

As shown in the plot, the estimated ORs of solving homicides comparing Black victims to White victims for most cities were relatively close and were smaller than 1, except that Fresno, CA and Stockton, CA had slightly higher ORs (between 1 and 1.5), and Albuquerque, NM had the highest OR (between 1.5 and 2). These three cities also had the largest range of confidence interval, indicating greater uncertainty. 

We can conclude that the odds of solving homicides with Black victims was smaller than the odds of solving homicides with White victims in most cities in the U.S..



## Problem 2
Load data and check for data types and any missing values
```{r}
birthweight_df = 
  read_csv("./data/birthweight.csv", na = c("", "NA", "Unknown"))

head(birthweight_df)
any(is.na(birthweight_df)) 
```
There is no missing data in this dataset.


Clean data for regression analysis (ensure proper data types)
```{r}
birthweight_df = 
  birthweight_df %>%
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  )
```


Propose a regression model for birthweight (based on a hypothesized structure for the factors that underly birthweight).
```{r}
# Based on my own prediction, I would love to study birthweight as an outcome that may depend on baby sex, father's and mother's race, and mother's weight gain during pregnancy.

proposed_model = 
  lm(bwt ~ babysex + mrace + frace + wtgain, data = birthweight_df)
```


Create a plot of model residuals against fitted values
```{r}
residuals_fitted_plot = 
  birthweight_df %>% 
  add_predictions(proposed_model) %>% 
  add_residuals(proposed_model) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Residuals vs. Fitted Values Plot"
  )

residuals_fitted_plot
```

According to the residual plot, the residuals are evenly distributed across fitted values, so we can conclude that the regression model follows homoscedasticity. 


Compare the proposed model to two others
```{r}
# One using length at birth and gestational age as predictors (main effects only)

# One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

cv_df = 
  crossv_mc(birthweight_df, 100) %>%
  mutate(
    proposed_mod = map(train, ~lm(bwt ~ babysex + mrace + frace + wtgain, data = birthweight_df)),
    main_mod = map(train, ~lm(bwt ~ blength + gaweeks, data = birthweight_df)),
    interaction_mod = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + blength*babysex + bhead*babysex + bhead*blength*babysex, data = .x))
  ) %>%
  mutate(
    rmse_proposed = map2_dbl(proposed_mod, test, ~rmse(model = .x, data = .y)),
    rmse_main = map2_dbl(main_mod, test, ~rmse(model = .x, data = .y)),
    rmse_interaction = map2_dbl(interaction_mod, test, ~rmse(model = .x, data = .y))
  )
```


Create plots to show the distribution of RMSE values for each model
```{r}
cv_df %>%
 select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(
    x = "Model",
    y = "RMSE",
    title = "Distribution of RMSE values for each model")
```

According to the plot, the RMSE value for proposed model is the highest, followed by the main effects model and then the interaction model. This indicates that the interaction model is the best fit model for the dataset among all three models.



## Problem 3
Download data for problem 3
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

head(weather_df)
```


Produce estimates of two quantities (r square, and log(b0*b1)) for each bootstrap sample
```{r}
boot_straps = 
  weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results_rsq = map(models, broom::glance),
    results_log = map(models, broom::tidy)
  ) %>%
  janitor::clean_names()

# Estimates of R-squared
rsq_df = 
  boot_straps %>% 
  select(-strap, -models) %>% 
  unnest(results_rsq) %>% 
  janitor::clean_names() %>% 
  select(id, r_squared)

# Estimates of log(b0*b1)
log_df = 
  boot_straps %>% 
  select(-strap, -models) %>% 
  unnest(results_log) %>% 
  select(id, term, estimate) %>%
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) %>% 
  janitor::clean_names() %>% 
  mutate(
    log_b0_b1 = log(intercept*tmin)
  )
```


Plot the distribution of two estimates and describe in words. 
```{r}
# Distribution plot of R-squared
rsq_dist = 
  rsq_df %>% 
  ggplot(aes(x = r_squared)) + 
  geom_density() +
  labs(
    title = "Distribution of R-squared",
    x = "R-squared",
    y = "Density"
  )

rsq_dist

# Distribution plot of log(b0*b1)
log_dist = 
  log_df %>% 
  ggplot(aes(x = log_b0_b1)) + 
  geom_density() +
  labs(
    title = "Distribution of log(b0*b1)",
    x = "log(b0*b1)",
    y = "Density"
  )

log_dist
```

As shown in the R-squared distribution plot, the curve is relatively symmetric, with a peak around R-squared of 0.913. As shown in the log(b0 times b1) distribution plot, the curve is slightly left-skewed, with a peak around log(b0 times b1) of 2.02.



Identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for R-squared and log(b0*b1)
```{r}
# For the R-squared
rsq_df %>%
  summarize(
    rsq_ci_lower = quantile(r_squared, 0.025), 
    rsq_ci_upper = quantile(r_squared, 0.975)
    ) %>% 
  knitr::kable(digit = 3)

# For the log(b0*b1)
log_df %>%
  summarize(
    log_ci_lower = quantile(log_b0_b1, 0.025), 
    log_ci_upper = quantile(log_b0_b1, 0.975)
    ) %>% 
  knitr::kable(digit = 3)
```

The 2.5% and 97.5% quantiles for R-squared are 0.894 and 0.927; the 95% confidence interval for R-squared is (0.894, 0.927). 

The 2.5% and 97.5% quantiles for log(b0 times b1) are 1.965 and 2.059; the 95% confidence interval for log(b0 times b1) is (1.965, 2.059).
